---
title: "Week 27: UFC Athletes and Fight Data"
date: 2026-07-07
description: "First challenge, a lot of setup and exploration.  Incomplete."
categories: [fightr, sport]
draft: false
execute:
  warning: false
  message: false
---

[Challenge Page on GitHub](https://github.com/rfordatascience/tidytuesday/tree/main/data/2026/2026-07-07)

## `fightr` Package

```{r}
#| eval: true

library(fightr)
```

Trying the code examples on the [fightr package GitHub page](https://github.com/benyamindsmith/fightr/)

```{r}
#| eval: true

# Loads from local cache, fetching updates automatically if needed
athletes <- get_ufc_data("ufc_athletes")
fights <- get_ufc_data("ufc_fights")

# Force a manual refresh of all datasets
update_all_ufc_data()
```


```{r}
#| eval: true

matchup_model <- model_fight_outcome(
    fighter1 = "Islam Makhachev",
    fighter2 = "Alexander Volkanovski",
    method_type = "bradley-terry"
)

# The function returns an S3 object containing the fitted model and predictions
print(matchup_model$method_probs)

# Extract and display the probability chart
matchup_model$plot
```


```{r}
#| eval: true

matchup_model <- model_fight_outcome(
    fighter1 = "Islam Makhachev",
    fighter2 = "Alexander Volkanovski",
    method_type = "multinomial"
)

# The function returns an S3 object containing the fitted model and predictions
print(matchup_model$method_probs)

# Extract and display the probability chart
matchup_model$plot
```


```{r}
#| eval: true

matchup_model <- model_fight_outcome(
    fighter1 = "Islam Makhachev",
    fighter2 = "Alexander Volkanovski",
    method_type = "thurstone-mosteller"
)

# The function returns an S3 object containing the fitted model and predictions
print(matchup_model$method_probs)

# Extract and display the probability chart
matchup_model$plot
```

### Probability Charts

The three probability charts shows that the probability that Alexander Volkanovski wins Islam Makhachev using KO/TKO is ~22%, using Submission is no more than 6%, and using Decision is ~20%.

### Technical Details

The function `get_ufc_data` exists in [`R/update_ufc_data.R`](https://github.com/benyamindsmith/fightr/blob/main/R/update_ufc_data.R).  The function reads data stored as `RData` from [`data/`](https://github.com/benyamindsmith/fightr/tree/main/data).  These files get downloaded and cleaned via a set of `Shell`, `R` and `Python` scripts stored in [`scripts/`](https://github.com/benyamindsmith/fightr/tree/main/scripts).  These scripts are executed periodically using GitHub workflow actions as specified in the `yml` files in [`.github/workflows/`](https://github.com/benyamindsmith/fightr/tree/main/.github/workflows).


## TidyTuesday Data

The challenge include 5 datasets.

```{r}
#| eval: true

local_file <- "data260707_cached.rds"

if(file.exists(local_file)) {
    message("Found local file! Loading cached data...")
    data260707 <- readRDS(local_file)
} else {
    message("Local file not found. Downloading from internet...")
    data260707 <- tidytuesdayR::tt_load('2026-07-07')
    saveRDS(data260707, file = local_file)
    message("Data downloaded and saved locally!")
}

data260707_small <- data260707$ufc_athletes
```

### EDA

```{r}
#| eval: true
#| output: asis

library(purrr)
library(inspectdf)

iwalk(data260707, function(df, name) {
    cat("\n\n####", name, "(", nrow(df), "," , ncol(df), ")\n")

    df |> inspect_types() |> show_plot() |> print()
    df |> inspect_cat() |> show_plot() |> print()
    df |> inspect_num() |> show_plot() |> print()
})
```

```{r}
#| eval: false

library(dplyr)
library(tidyselect)
library(purrr)
library(GGally)

iwalk(data260707, function(df, name) {
    cat("####", name, row(df), "," , col(df), "\n")

    df_safe <- df |> 
        select(where(~ is.numeric(.) || (is.character(.) && n_distinct(.) <= 15)))

    ggpairs(df_safe) |> print()    
})
```


```{r}
#| eval: false
#| results: asis

library(summarytools)
library(purrr)

iwalk(data260707, function(df, name) {
  cat("####", name, row(df), "," , col(df), "\n")


  dfSummary(
    df,
    plain.ascii = FALSE,
    style = "grid",
    graph.magnif = 0.82,
    varnumbers = FALSE,
    valid.col = FALSE,
    temp.img.dir = "/tmp",
    round.digits = 2,
    headings = FALSE,
    display.labels = FALSE
  ) |>
    print()
})
```