---
title: "Week 27: UFC Athletes and Fight Data"
date: 2026-07-07
description: "Setup + explore EDA options"
categories: [fightr, sport]
draft: false
---

```{r}
#| include: false

week = rmarkdown::metadata$date
week = "2026-07-07"
```

```{r}
#| include: false
#| eval: true

source("../../scripts/md.R")
```

```{r}
#| output: asis
#| echo: false

generateTtLink(week)
```


## Explore `fightr` Package

```{r}
#| include: true
#| eval: true

library(fightr)
```

Trying [package](https://github.com/benyamindsmith/fightr/) code examples.

Read data
```{r}
#| include: true
#| eval: true

# Loads from local cache, fetching updates automatically if needed
athletes <- get_ufc_data("ufc_athletes")
fights <- get_ufc_data("ufc_fights")

# Force a manual refresh of all datasets
update_all_ufc_data()
```

Estimation Method 1: `bradley-terry`
```{r}
#| include: true
#| eval: true

matchup_model <- model_fight_outcome(
    fighter1 = "Islam Makhachev",
    fighter2 = "Alexander Volkanovski",
    method_type = "bradley-terry"
)

# The function returns an S3 object containing the fitted model and predictions
print(matchup_model$method_probs)

# Extract and display the probability chart
matchup_model$plot
```

Estimation Method 2: `multinomial`
```{r}
#| include: true
#| eval: true

matchup_model <- model_fight_outcome(
    fighter1 = "Islam Makhachev",
    fighter2 = "Alexander Volkanovski",
    method_type = "multinomial"
)

# The function returns an S3 object containing the fitted model and predictions
print(matchup_model$method_probs)

# Extract and display the probability chart
matchup_model$plot
```

Estimation Method 2: `thurstone-mosteller`
```{r}
#| include: true
#| eval: true

matchup_model <- model_fight_outcome(
    fighter1 = "Islam Makhachev",
    fighter2 = "Alexander Volkanovski",
    method_type = "thurstone-mosteller"
)

# The function returns an S3 object containing the fitted model and predictions
print(matchup_model$method_probs)

# Extract and display the probability chart
matchup_model$plot
```

**Probability Charts.** The three probability charts shows that the probability that Alexander Volkanovski wins Islam Makhachev using KO/TKO is ~22%, using Submission is no more than 6%, and using Decision is ~20%.

**Technical Details.**  The function `get_ufc_data` exists in [`R/update_ufc_data.R`](https://github.com/benyamindsmith/fightr/blob/main/R/update_ufc_data.R).  The function reads data stored as `RData` from [`data/`](https://github.com/benyamindsmith/fightr/tree/main/data).  These files get downloaded and cleaned via a set of `Shell`, `R` and `Python` scripts stored in [`scripts/`](https://github.com/benyamindsmith/fightr/tree/main/scripts).  These scripts are executed periodically using GitHub workflow actions as specified in the `yml` files in [`.github/workflows/`](https://github.com/benyamindsmith/fightr/tree/main/.github/workflows).


## EDA

```{r}
#| include: false
#| eval: true

source("../../scripts/qread.R")
data <- qread(week = week)
```

```{r}
#| echo: false
#| eval: true
#| results: asis

source("../../scripts/viz.R")
viz(data = data)
```
